Equation-free analysis of a dynamically evolving multigraph

### 2016-04-27 manuscript-materials/tex/response.tex

\documentclass[12pt]{article}

\usepackage{graphicx}
%\usepackage{subcaption}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{epstopdf}
%\usepackage{fullpage}

\begin{document}

\begin{flushright}
From: A. Holiday, I. Kevrekidis
\end{flushright}

\vspace{5mm}

% \begin{flushleft}
% To: Terre Arena \\
% Senior Editorial Assistant \\
% Physical Review E \\
% Email: pre@aps.org \\
% http://journals.aps.org/pre/
% \end{flushleft}

\vspace{2mm}

\section*{Response to comments/changes implemented in Equation-free analysis
of a dynamically evolving multigraph by A. Holiday and I. Kevrekidis.}

First, we would like to thank both reviewers for their helpful input!

\vspace{3mm}

\textbf{First Editor's recommendations}
\vspace*{\baselineskip}

* {\em ... I would like to make a few suggestions to the authors of
  the paper for a possible extension of their methodology in the
  investigation of more realistic models of complex networks and
  possibly models supported by concrete empirical data. The notion of
  a multigraph is too general to include the whole range of
  applications of graph theory in complex network science. As a matter
  of fact, a further step would be towards the consideration of
  attributes (or labels) on nodes and links of a (multi)graph... }.

\textbf{OUR REPLY}: It is a very good suggestion to apply the method
to other systems with labeled nodes, and we have in fact already
addressed a first example of this in \cite{kattis_modeling_2015} in which
an SIS epidemiological model is investigated with similar
techniques. We have added a reference to this paper in the text.
\vspace*{\baselineskip}

* {\em ... in the literature of network science there are many
  examples of such large empirical networks exhibiting structures of a
  multilayer (or multiplex) typology. Equation-free modeling applied
  to such empirically sustained networks would be expected to
  demonstrate a very interesting novel attainment in the evolving
  field of complex network science.}

\textbf{OUR REPLY}: An application to empirically-based network models
would also be worthwhile, but as the aim of the current paper is
simply to show the extension of our techniques to systems with
multiple edges, we believe the recommendation would serve well as a
separate, second paper on the topic which we are already working on.
\vspace*{\baselineskip}

\textbf{Second Editor's recommendations}

* {\em While it is clear that the paper deals with new techniques
  applied in well known systems, it is not clear what questions it
  tries to answer, what are the objectives of this work. Thus, I would
  try to add a couple of sentences both in the abstract and in the
  conclusion sections showing where such results are useful, other
  that as theoretical techniques.}

\textbf{OUR REPLY}: To clarify the prospects for future application
areas, we have added a little information in the abstract and
conclusion section mentioning relevant systems that could benefit from
such a study.  \vspace*{\baselineskip}

{\em There is a Figure 1, but it is not discussed in the text of the
  paper.  Figure 2. It is not clear which is the independent
  variable. Is it the degree or is it the vertex number? Are the
  numbers given here summations of many simulations?  How many?}

\textbf{OUR REPLY}: We have added a missing reference to Fig. 1, and
appreciate the reviewer's help in noticing its absence. To clarify
Fig. 2 we have created a new caption. Indeed the results are an
instantaneous average over twenty simulations. Also, the vertex number
is simply a pseudo-labeling of each node for the purpose of
plotting. In the actual model, vertices are not labeled.
\vspace*{\baselineskip}

{\em It is clear here that multiple links are allowed between two
  specific nodes.  While this may be a mathematical notion, all
  practical problems in networks today deal with a single link between
  two nodes. Would the results be the same for such a case?}

\textbf{OUR REPLY}: The model could also be cast as a weighted
network, in which the edge weight signifies the number of connections
between two nodes. We now note this in the model description. While
this might be a more familiar formulation to some researchers, since
previous theoretical results on the system were couched in the
language of multiple edges, we chose this framework. 
\vspace*{\baselineskip}

\bibliographystyle{abbrv} \bibliography{multigraph-refs.bib}

\end{document}

### 2016-04-27 manuscript-materials/tex/multigraph-manuscript.tex
Small changes to manuscript, added respone to reviewers

\documentclass[epjST, final]{svjour} \usepackage{graphicx, subcaption,
  setspace, indentfirst, amsfonts, amsthm, empheq}
% \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
% \usepackage[center]{titlesec}
\graphicspath{ {../figs/} }
% \pagestyle{plain} \renewcommand\thesection{\Roman{section}}
% \renewcommand\thesubsection{\alph{subsection}.}
\newcommand\ignore[1]{}
% \setlength{\parskip}{0ex}
% \def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
% \let\endchangemargin=\endlist

@@

\abstract{In order to illustrate the adaptation of traditional
  continuum numerical techniques to the study of complex network
  systems, we use the equation-free framework to analyze a dynamically
  evolving multigraph.
%
  This approach is based on coupling short intervals of direct dynamic
  network simulation with appropriately-defined lifting and
  restriction operators, mapping the detailed network description to
  suitable macroscopic (coarse-grained) variables and back.
%
  This enables the acceleration of direct simulations through Coarse
  Projective Integration (CPI), as well as the identification of
  coarse stationary states via a Newton-GMRES method.
%
  We also demonstrate the use of data-mining, both linear (principal
  component analysis, PCA) and nonlinear (diffusion maps, DMAPS) to
  determine good macroscopic variables (observables) through which one
  can coarse-grain the model.}

@@

  \noindent where $d_i$ is the degree of vertex $v_i$, $E(G)$ is the
  set of edges in the graph, and $\kappa \in (0, \infty)$ is a model
  parameter affecting the influence degrees have on the probability
  distribution in Step \ref{step:prob}.
%
  That is, taking $\lim\limits{\kappa \rightarrow 0}$, we recover
  ``pure'' preferential attachment, and probability scales directly
  with degree,
  %%% YGK immediately? after a short time ? at steady state ?  if at
  %%% steady state say it ?  ATH This refers to the dynamics at all
  %%% times. At every step, we need some probability of choosing a
  %%% given vertex v_i: if k -> 0, this probability is proportional to
  %%% deg(v_i), and if k -> infinity this probability is uniform over
  %%% all vertices (and degree has no effect. Valid at every step.
  while
  $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n}
  \; \forall \; l$, and individual degrees have no effect.
  %%% YGK you mean individual initial degrees ?
%%%
  \par

@@

  \begin{figure}[th!]
    \vspace{-5mm} \centering
    \begin{subfigure}[t]{0.49\textwidth}
      \centering
      \includegraphics[width=\textwidth]{lopsided-degs-a}
      \subcaption{\label{fig:lopsided-init}}
    \end{subfigure} %
    \begin{subfigure}[t]{0.49\textwidth}
      \centering
      \includegraphics[width=\textwidth]{erdos-degs-a}
      \subcaption{\label{fig:erdos-init}}
    \end{subfigure}
    \caption{Multigraph evolution dynamics for an $n=100$ vertex graph
      observed through the degree sequence (after sorting the nodes by
      degree). Two distinct transients are shown, initialized with (a)
      an Erd\H{o}s-R\'{e}nyi random graph with $m = 5,050$ edges and (b)
      a graph in which half the vertices are isolated, and half
      uniformly share $m = 5,050$ edges. Both approach the same ultimate
      stationary sequence. \label{fig:dse}}
  \end{figure}

@@

  \subsection{Coarse-projective integration (CPI)}
  \label{sec:cpi}

@@

  The equation free framework was successfully applied to the
  edge-conservative preferential attachment model, accelerating
  simulations through CPI and locating stationary states through coarse
  Newton-GMRES.
  % 
  Additionally, an underlying two-dimensional description was uncovered via PCA and DMAPS. \\

  An open area of investigation is the interpretation of the output from
  the PCA and DMAPS techniques.
  % 
  As a linear map, PCA retains a clear relationship between its input
  and output, but is less effective when data lie on highly nonlinear
  manifolds.
  % 
  DMAPS may perform better the task of dimensionality reduction, but it
  is unclear how the embedding coordinates relate to physical features
  of the sampled networks.
  % 
  While this approach opens several promising avenues in coarse-graining
  complex network dynamics, it also reveals two ``bottlenecks'' for the
  process: (a) the selection of informative and practically computable
  graph distance metrics, necessary in the discovery of good coarse
  variables; and (b) the construction of network realizations consistent
  with (conditioned on) specific values of collective observables.
  % 
  Both items are, and we expect will continue being, the subject of
  intense investigation by many research groups, including ours.

>>>>

\documentclass[epjST, final]{svjour} \usepackage{graphicx, subcaption,
  setspace, indentfirst, amsfonts, amsthm, empheq, color, soul}
% \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
% \usepackage[center]{titlesec}
\graphicspath{ {../figs/} }
% \pagestyle{plain} \renewcommand\thesection{\Roman{section}}
% \renewcommand\thesubsection{\alph{subsection}.}
\newcommand\ignore[1]{}
% \setlength{\parskip}{0ex}
% \def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
% \let\endchangemargin=\endlist

@@

\abstract{In order to illustrate the adaptation of traditional
  continuum numerical techniques to the study of complex network
  systems, we use the equation-free framework to analyze a dynamically
  evolving multigraph.
%
  This approach is based on coupling short intervals of direct dynamic
  network simulation with appropriately-defined lifting and
  restriction operators, mapping the detailed network description to
  suitable macroscopic (coarse-grained) variables and back.
%
  This enables the acceleration of direct simulations through Coarse
  Projective Integration (CPI), as well as the identification of
  coarse stationary states via a Newton-GMRES method.
%
  We also demonstrate the use of data-mining, both linear (principal
  component analysis, PCA) and nonlinear (diffusion maps, DMAPS) to
  determine good macroscopic variables (observables) through which one
  can coarse-grain the model.
%
  These results suggest methods for decreasing simulation times of
  dynamic real-world systems such as epidemiological network
  models. Additionally, the data-mining techniques could be applied to
  a diverse class of problems to search for a succint, low-dimensional
  description of the system in a small number of variables.}

@@

  \noindent where $d_i$ is the degree of vertex $v_i$, $E(G)$ is the
  set of edges in the graph, and $\kappa \in (0, \infty)$ is a model
  parameter affecting the influence degrees have on the probability
  distribution in Step \ref{step:prob}. 
%
  That is, taking $\lim\limits{\kappa \rightarrow 0}$, we recover
  ``pure'' preferential attachment, and probability scales directly
  with degree,
  %%% YGK immediately? after a short time ? at steady state ?  if at
  %%% steady state say it ?  ATH This refers to the dynamics at all
  %%% times. At every step, we need some probability of choosing a
  %%% given vertex v_i: if k -> 0, this probability is proportional to
  %%% deg(v_i), and if k -> infinity this probability is uniform over
  %%% all vertices (and degree has no effect. Valid at every step.
  while
  $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n}
  \; \forall \; l$, and individual degrees have no effect. A single
  step of this evolution process is illustrated in
  Fig. ({\ref{fig:step-illustration}}). Note that this can also be
  represented as a graph in which only one edge is permitted between
  each pair of nodes, but with edge weights signifying the number, or
  strength, of connections between nodes.
  %%% YGK you mean individual initial degrees ?
%%%
  \par

@@

  \begin{figure}[th!]
    \vspace{-5mm} \centering
    \begin{subfigure}[t]{0.49\textwidth}
      \centering
      \includegraphics[width=\textwidth]{lopsided-degs-a}
      \subcaption{\label{fig:lopsided-init}}
    \end{subfigure} %
    \begin{subfigure}[t]{0.49\textwidth}
      \centering
      \includegraphics[width=\textwidth]{erdos-degs-a}
      \subcaption{\label{fig:erdos-init}}
    \end{subfigure}
    \caption{Evolution of the multigraph model's sorted degree
      sequence. Two distinct transients are shown, initialized with
      (a) an Erd\H{o}s-R\'{e}nyi random graph with $m = 5,050$ edges
      and (b) a graph in which half the vertices are isolated, and
      half uniformly share $m = 5,050$ edges. Both approach the same
      ultimate stationary sequence. To smoothen the evolution of this
      stochastic system, here we plot the instantaneous average of twenty
      simulations. \label{fig:dse}}
    % Multigraph evolution dynamics for an $n=100$ vertex graph
    % observed through the degree sequence (after sorting the nodes by
    % degree)
  \end{figure}

@@

  \subsection{Coarse projective integration (CPI)}
  \label{sec:cpi}

@@

  The equation free framework was successfully applied to the
  edge-conservative preferential attachment model, accelerating
  simulations through CPI and locating stationary states through coarse
  Newton-GMRES. This indicates potential avenues for improving
  simulation times of other complex network models, such as those
  found in epidemiology.
  % 
  Additionally, an underlying two-dimensional description was
  uncovered via PCA and DMAPS. These automated methods of
  dimensionality-reduction are quite general, and can in principle be
  applied in a wide range of network settings to probe hidden
  low-dimensional structure. For an application in the setting of
  labeled nodes, see ({\cite{kattis_modeling_2015}}). \\

  However, an open area of investigation is the interpretation of the output from
  the PCA and DMAPS techniques.
  % 
  As a linear map, PCA retains a clear relationship between its input
  and output, but is less effective when data lie on highly nonlinear
  manifolds.
  % 
  DMAPS may perform better the task of dimensionality reduction, but it
  is unclear how the embedding coordinates relate to physical features
  of the sampled networks.
  % 
  While this approach opens several promising avenues in coarse-graining
  complex network dynamics, it also reveals two ``bottlenecks'' for the
  process: (a) the selection of informative and practically computable
  graph distance metrics, necessary in the discovery of good coarse
  variables; and (b) the construction of network realizations consistent
  with (conditioned on) specific values of collective observables.
  % 
  Both items are, and we expect will continue being, the subject of
  intense investigation by many research groups, including ours.
